A National Pipeline System as a Solution to Future Water Shortage in the United States

### 2013-02-04 paper.tex
added 25minus05

@@

Edges in the Minimal Spanning Tree in Figure \ref{fig:graph} are weighted inversely proportional to cost and directly proportional to the populations at each of the endpoints, and sorted. The resulting sorted table gives a full list of all $3108$ pipelines sorted by this utility metric. The final (highest priority) rows are shown in Table \ref{table:priority}.

\begin{table}[h]\label{table:priority}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
FIPS 1 & FIPS 2 & Priority \tabularnewline
\hline
48201 & 48339 & 10346.3 \tabularnewline
\hline
24005 & 24510 & 10675.4 \tabularnewline
\hline
6001  &6013 & 11019.8 \tabularnewline
\hline
34003 & 34017 & 11190.7 \tabularnewline
\hline
51153 & 51685 & 12104.5 \tabularnewline
\hline
11001 & 51013 & 12113.6 \tabularnewline
\hline
48029 & 48091 & 13787.1 \tabularnewline
\hline
34017 & 36061 & 16584.3 \tabularnewline
\hline
51059 & 51600 & 18528.8 \tabularnewline
\hline
51153 & 51683 & 22134.2 \tabularnewline
\hline
6037 & 6059  &23742.4 \tabularnewline
\hline
36047 & 36061  &27446.1 \tabularnewline
\hline
36059 & 36081  &29256.5 \tabularnewline
\hline
36005 & 36061 & 32595 \tabularnewline
\hline
36047 & 36081 & 33517.6 \tabularnewline
\hline
17031 & 17043 & 41637.4 \tabularnewline
\hline

@@

\subsection*{Computational Software}
All data analysis and visualization was performed using the technical computing language Julia \cite{julia}, making use of the packages "DataFrames", "Gadfly", "Distributions", "Compose", "Graphs" all of which are available in the Julia package manager and were licensed under the MIT license.

@@

\subsubsection{Data Processing}
The dataset we obtained was in the proprietary, binary "Shapefile" format, that is used by many professional GIS systems. Luckily, there exists an open documentation of the format \cite{shapefile-gis}, which allowed us to implement our own parser to extract the locations, and shapes of the various counties.

@@

As outlined in Section 3, we made use various data points collected by the USGS (United States Geological Survey) for every single county in the United  States in the years 1985, 1990, 1995, 2000 and 2005. The data is split into 104 categories ranging from standard data points such as the total amount of water consumed in a specific specific county to more unexpected ones such as water irrigation due to golf courses. In addition to the actual data, every row also includes meta information such as the State in which the county is located and the FIPS (Federal Information Processing Standard) code of the county for which the data was collected.

@@

\subsubsection{Data Processing}
The standard format of the data allowed for an easy import into our computational software using the Julia "DataFrames" package. This allowed us to treat the data essentially as an im-memory database which we could query not only for the downloaded data, but which also allowed for more complex models to be executed on the data with relative ease. Treating the data as a database also tremendously helped with combining all the data from the many different sources. Not all of the data we obtained contained the same information as to their location, but after normalizing all the different kinds of information available to a FIPS code based primary key, standard database operations such as joining two datasets by primary key. This operation was incredibly helpful during visualization as we could just join the DataFrame containing the geographical information with that containing e.g. the water usage information and have the results be immediately

@@

        \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{95.png}
                \caption{1995}
                \label{fig:1995}
        \end{subfigure}
       \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{00.png}
                \caption{2000}
                \label{fig:2000}
        \end{subfigure}
        \caption{USGS Data on total water consumption in the indicated year. Values are represent on a logarithmic scale with green indicating values towards the bottom of the scale and red indicating values towards the top of the scale.}\label{fig:water}
\end{figure}

In order to be able to visualize the input data and any simulation results obtained from the models used in this paper, we created a custom visualization software based on the Julia "Compose" package, a modern implementation of Wilkinson-Wickham's Grammar of Graphics.

The visualization software can visualize arbitrary county level data sets as long as there is some way to identify  which county the data belongs to. This is easiest is the data in need of visualization is already keyed on the FIPS code of the county, but it is also possible to plot data that only contains longitudinal and latitudinal information of any point in the county, a feature which is very useful to visualize data coming from individual measurement stations within the county rather than values that are determined for the county as a whole.

We extended our software to be able to graph arbitrary directed and undirected Graphs (represented by the Julia "Graphs") between counties our points within these counties. This capability was used to generate the graph of the proposed pipeline network.

Our software also makes use of the vector graphics capabilities of the Compose framework to create almost infinitely scalable graphics that allowed us to visualize smaller areas very effectively. The figures contained within this document are rasterizations of these vector images  to ensure a reasonable file size.

@@

While de-salinization is simply not an option at all for much of central
America, one idea we considered was the construction of additional
dams and reservoirs. One considerable environmental issue associated
with large dams is that they can cause unanticipated changes in overall
river flow\cite{key-17}, totaling potentially huge amounts in damages.

>>>>

\begin{figure}[h]
        \centering
        \includegraphics[width=\textwidth]{25minus05}
        \caption{Visualization of projected increase in water usage, by county. Red is high, green is low, and cyan is projected decrease.}
        \label{fig:outline}
\end{figure}

@@

Edges in the Minimal Spanning Tree in Figure \ref{fig:graph} are weighted inversely proportional to cost and directly proportional to the populations at each of the endpoints, and sorted. The resulting sorted table gives a full list of all $3108$ pipelines sorted by this utility metric. The final (highest priority) rows are shown in Table 2.

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
FIPS 1 & FIPS 2 & Priority \tabularnewline
\hline
48201 & 48339 & 10346.3 \tabularnewline
\hline
24005 & 24510 & 10675.4 \tabularnewline
\hline
6001  &6013 & 11019.8 \tabularnewline
\hline
34003 & 34017 & 11190.7 \tabularnewline
\hline
51153 & 51685 & 12104.5 \tabularnewline
\hline
11001 & 51013 & 12113.6 \tabularnewline
\hline
48029 & 48091 & 13787.1 \tabularnewline
\hline
34017 & 36061 & 16584.3 \tabularnewline
\hline
51059 & 51600 & 18528.8 \tabularnewline
\hline
51153 & 51683 & 22134.2 \tabularnewline
\hline
6037 & 6059  &23742.4 \tabularnewline
\hline
36047 & 36061  &27446.1 \tabularnewline
\hline
36059 & 36081  &29256.5 \tabularnewline
\hline
36005 & 36061 & 32595 \tabularnewline
\hline
36047 & 36081 & 33517.6 \tabularnewline
\hline
17031 & 17043 & 41637.4 \tabularnewline
\hline

@@

\subsection*{Computational Software}
All data analysis and visualization was performed using the technical computing language Julia, making use of the packages "DataFrames", "Gadfly", "Distributions", "Compose", "Graphs" all of which are available in the Julia package manager and were licensed under the MIT license.

@@

\subsubsection{Data Processing}
The dataset we obtained was in the proprietary, binary "Shapefile" format, that is used by many professional GIS systems. Luckily, there exists an open documentation of the format \cite{key-21}, which allowed us to implement our own parser to extract the locations, and shapes of the various counties.

@@

As outlined in Section 3, we made use various data points collected by the USGS (United States Geological Survey) for every single county in the United  States in the years 1985, 1990, 1995, 2000 and 2005. The data is split into 104 categories ranging from standard data points such as the total amount of water consumed in a specific specific county to more unexpected ones such as water irrigation due to golf courses.

@@

@@

        \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{95.png}
                \caption{1995}
                \label{fig:1995}
        \end{subfigure}
       \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{00.png}
                \caption{2000}
                \label{fig:2000}
        \end{subfigure}
        \caption{USGS Data on total water consumption in the indicated year. The more red, the higher the water consumption of the county.}\label{fig:water}
\end{figure}

For this data, we wrote a visualization software that allows arbitrary sets of county-level data to be displayed on maps of the United States. The visualization of total water usage for every county in the years covered by the USGS data is shown in figure \ref{fig:water}

@@

While de-salinization is simply not an option at all for much of central
America, one idea we considered was the construction of additional
dams and reservoirs. One considerable environmental issue associated
with large dams is that they can cause unanticipated changes in overall
river flow \cite{key-17}, totaling potentially huge amounts in damages.
